---
title: "Assignment 4"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Data

```{r}
#Import Data Set
data <- read.csv("/Users/sivandubinsky/Documents/Uni/ST362 Master/ST362-Projects/Assignment 4/Medical_insurance.csv")
```
# Exploratory Data Analysis

Here, I am graphing the pairs of different numerical variables and trying to see if there is any clear correlation between any two variables. 

```{r}
pairs(data[c("age","bmi","children","charges")])
```
The two noticible paairs are bmi/charges and age/charges. As it seems like there are cohorts within those observations. Let's analyze bmi and charges and see if smoking is the differentiator. Let's also do the same for age

```{r}
plot(data$bmi,data$charges)
plot(data$age,data$charges)
```
Here we can infer that age is not the major factor when determining medical charges. This is because people with the same age are broken down into three different charge levels. This means that there is some other major factor. Similarly, looking at the bmi/charges plot we can see two things.

1. Charges seem to start increasing past a 30 bmi. This is also characterized as being obese
2. There is a cohort of individuals who are consistently being charged more then the others. Having some epxerience in the insurance space, my hunch is that this has something to do with smokers. Lets  explore further and see whether being a smoker has a singificant impact on medical charges

```{r}
boxplot(charges ~ smoker, data = data,
        xlab = "Smoker", ylab = "Charges",
        main = "Charges Boxplot by Smoke")
```
Here we can see that in fact, smokers pay more in medical fees then non-medical fees. Lets look at our previous two plots again but this time we'll colour the points blue if the individual is a non-smoker and red if they are a smoker

```{r}
colours <- ifelse(data$smoker == "yes", "red", "blue")

plot(data$bmi,data$charges,col=colours)
plot(data$age,data$charges,col=colours)
```
From the data we can see that smoking has a significant impact on medical charges. Notice, that while smokers tend to have higher medical charges, smokers who are obese (bmi >= 30) have significantly higher medical charges. This is something we should keep in mind as we further analyze the data. Further, when comparing age to charges, it appears that the cohort made up entirely of smokers, by far has the highest medical expenses. This once again reassures me of the fact that smoking is a very important factor for determining medical charges. 


#Applying a model

First, lets look at a model with all predictors and see if the p-values can tell us anything about the significance of certain predictors

```{r}
model1 <- lm(charges~.,data=data)
summary(model1)
```
Here we see that 

- Sex is not having a significant impact on the model
- The fact that someone lives in the northwest region does not have a significant impact. The other regions however do have some significance. 

```{r}

#A person is considered obese if their BMI is over 30. Lets create a new column that tells us if a person is obese or not

data$obese <- ifelse(data$bmi >= 30, 1, 0)

data

``` 


```{r}
#New model with obesity as a categorical variable, sex dropped
model2<- lm(charges~age+obese+children+smoker+SW_SE,data=data)
summary(model2)
```
Another assumption I would like to check is whether or not age should be polynomial, specifically quadratic This is because from personal experience, as age increases, the size of medical charges tends to go up as health problems become more serious. Further, looking at the data it is possible to say that there is a quadratic relationship between age and charges but it is not clear. Let's lets do a significance test and see what the relationship should be with charges.

```{r}

ggplot(data, aes(x = predict(lm(charges~age),data=data), y = charges)) + 
  geom_point(color = "red", alpha = 0.7) + 
  geom_abline(color = "black") +
  ggtitle("Prediction vs. Actual Values")

ggplot(data, aes(x=age, y=charges)) + 
          geom_point() +
          stat_smooth(method='lm', formula = y ~ poly(x,2), size = 1) + 
          xlab('Age') +
          ylab('Charges')

```

```{r}


```



```{r}
#Consider that when people start to age, it is likely that their medical expense will go-up quicker than when they were younger.
#Lets incorporate this into our model

model3<- lm(charges~poly(age,2) + bmi + children + smoker + smoker*obese + region, data=data)
summary(model3)

library(ggplot2)

ggplot(data, aes(x = predict(model3,data=data), y = charges)) + 
  geom_point(color = "red", alpha = 0.7) + 
  geom_abline(color = "black") +
  ggtitle("Prediction vs. Actual Values")
```
## Final Model

$$y = age^2 + bmi + children + smoker + region + smoker*obese$$

